Tree-based

### 2023-12-15 EDA.Rmd
Added comparison to variables in unspervised
idk

# EDA

## Exploring the demographics of the patients

@@

## Looking at the data itself

@@

An observation is that many of the variables have 106 NAs. This indicates that a good fraction [106 patients] were equally tracked and measured in a less extensive way.

@@

## Exploring some interactions of variables

@@

### mihir choose your preffered graph and explain

@@

This question investigates the effectiveness of the drug D-penicillamine compared to a placebo, considering the stage of PBC. By analyzing survival rates across different disease stages and treatment types, we can assess the drug's effectiveness at various disease stages. This information is vital for clinicians to make informed decisions about treatment plans and for researchers to understand the drug's impact. (LOLL this is what the study acc sought to discover)

>>>>

# Exploratory Data Analysis

## Part 1: Exploring the demographics of the patients

@@

## Part 2: Looking at the data itself

@@

An observation is that many of the variables have 106 NAs. This indicates that a good fraction [106 patients] could have been equally tracked and measured in a less extensive way.

@@

## Part 3: Exploring some interactions of variables

@@

@@

This question investigates the effectiveness of the drug D-penicillamine compared to a placebo, considering the stage of PBC. By analyzing survival rates across different disease stages and treatment types, we can assess the drug's effectiveness at various disease stages. This information is vital for clinicians to make informed decisions about treatment plans and for researchers to understand the drug's impact.

### 2023-12-15 Regression.Rmd
Wrote the introduction for the data
idk

This analysis suggests significant associations between several predictors (Bilirubin, Age, Alk_Phos, Prothrombin, Spiders) and the status of the patient. Each predictor, including the Intercept, shows a statistically significant effect, as indicated by their small p-values (Pr(\>\|z\|)). The negative coefficients for Bilirubin, Age, Alk_Phos, Prothrombin, and Spiders imply that increases in these variables are associated with a decrease in the log-odds of the outcome. Bilirubin and Spiders are showing notably strong effects in the model.

@@

This figure shows us the Tukey-Anscombe plot using Pearson residuals for Model 2. In an ideal fit, the residuals should be evenly distributed about zero with constant mean and variance. The shape of the line suggests that the model is not capturing someunder lying structure in the data, especially in extreme cases.

@@

Lastly, let us look at the ROC curves for bothe the models.

@@

Model 1: - Its ROC curve is above the line of no discrimination, indicating that the model has some predictive capabilities. - The AUC is 0.64, which is better than random guessing but suggests there's room for improvement.

Model 2: - The ROC curve for Model 2 is significantly above that of Model 1, and much closer to the top-left corner, indicating better predictive performance. - The AUC is 0.899, which suggests a good classification performance, and it's notably better than Model 1. - Its ability to discriminate between positive and negative classes is superior to that of Model 1.

@@

Using our final model, we now turn to see how well this classifier does on our testing data. Recall that we initally set the test data during the train test split.

@@

This shows that our model is has struggled with overfitting, with an especially low precision score. We will discuss this in the comparison with the tree model, but this is a limitation of our data size. Lastly, we can plot the separation plot.

>>>>

This is the role of inference in evaluating our model. The most notable predictors are Bilirubin, Age, Alk_Phos, and Prothrombin, each showing a statistically significant relationship (p \< 0.05) with the dependent variable. The Intercept and EdemaY have extremely significant p-values, but the practical significance of EdemaY is questionable due to its large standard error. Other variables like Spiders, SGOT, DrugPlacebo, and Copper, while contributing to the model, do not reach conventional levels of statistical significance (p \< 0.05).

@@

This figure shows us the Tukey-Anscombe plot using Pearson residuals for Model 2. In an ideal fit, the residuals should be evenly distributed about zero with constant mean and variance. The shape of the line suggests that the model is not capturing someunder lying structure in the datain extreme cases.

@@

Lastly, let us look at the ROC curves for both the models.

@@

Model 1:

-   Its ROC curve is above the line of no discrimination, indicating that the model has some predictive capabilities

-   The AUC is 0.638, which is better than random guessing but suggests there's room for improvement.

Model 2:

-   The ROC curve for Model 2 is significantly above that of Model 1, and much closer to the top-left corner, indicating better predictive performance.

-   The AUC is 0.899, which suggests a good classification performance, and it's notably better than Model 1.

-   Its ability to discriminate between positive and negative classes is superior to that of Model 1.

@@

Using our final model, we now turn to see how well this classifier does on our testing data. Recall that we initially set the test data during the train test split.

@@

This shows that our model is has struggled with overfitting, with an especially low precision score. We will discuss this in the comparison with the tree model, but this is an importnat limitation of our data size as we discussed in the introduction. Lastly, we can plot the separation plot. We can see the separation has increased due to the over fitting we discussed.

### 2023-12-15 Tree.Rmd
Wrote the introduction for the data
idk

# Part I: Classification Algorithm for the Status of the Patient

@@

In this report, I have decided to use {rpart} because it provides more flexibility for surrogate splits and the trees are a bit easier to make attractive looking.

@@

(Explanation of above) To gain a further understanding of the data, we can plot a tree yielding Collection Node style trees. This can help us understand how the data is split.

@@

@@

From the graph we can see that a cp of 0.011 is ideal as it is under the horizontal (dotted) reference line. We can prune the tree with this CP value.

@@

@@

kable(
  conf_matrix,
  col.names = c("Prediction/Supervision", "0", "1"),
  digits = 3,
  booktabs = TRUE,
  caption = "Model 1: Confusion Matrix",
  align = "c"
) %>%
kable_styling(latex_options = "HOLD_position")

@@

From this, we can get the model's

-   True Positive (TP): Predicted Graduated and actually Graduated. In this case, 2217

-   False Positive (FP): Predicted Graduated but actually Not Graduated. In this case, 1000

-   False Negative (FN): Predicted Not Graduated but actually Graduated. In this case, 244

-   True Negative (TN): Predicted Not Graduated and actually Not Graduated. In this case, 659

With this, we can also calculate the model's metrics.

@@

>>>>

# Part I: Classification Algorithm for the Status of the Patient using CART

@@

In this report, we have decided to use {rpart} because it provides more flexibility for surrogate splits and the trees are a bit easier to make attractive looking.

@@

To gain a further understanding of the data, we can plot a tree yielding Collection Node style trees. This can help us understand how the data is split.

@@

The tree shows us the splits that were done on Age, Bilirubin and Prothrombin. Interestingly, Stage did not contribute in the tree.

@@

From the graph we can see that a cp of 0.039 is ideal as it is under the horizontal (dotted) reference line. We can prune the tree with this CP value.

@@

We ca see the pruned tree has cut out some leaf nodes. This would help in avoiding overfitting the model.

@@

kable(
  conf_matrix,
  col.names = c("Prediction/Supervision", "0", "1"),
  digits = 3,
  booktabs = TRUE,
  caption = "Model 1: Confusion Matrix (0=Deceased, 1=Censored)",
  align = "c"
) %>%
kable_styling(latex_options = "HOLD_position")

@@

With this, we can also calculate the model's metrics on the test data.

@@

As you can see the model shows good accuracy with 72.9%. However, the specificity is an issue with 57.6%. Now let us compare this with a different type of model: logistic regression.
